Global patterns of forest autotrophic carbon fluxes

### 2020-06-13 manuscript/manuscript draft.Rmd
finished hm comments on manuscript pdf
These are all now dealt with or covered in issues.

Closes #81

All FACF increased linearly with MAT, and we found no support for a saturation point of FACF with MAT (all p<0.05; Figs. 3-4, S4-S5, Table S2). As with latitude, MAT tended to explain more variation in the larger FACF ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).

MAP was a significant (p<0.05) predictor of all FACF except $ANPP_{stem}$ (Figs. 4a, S4-S5; Table S2). However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 37% of variation in FACF. For the majority of FACF, a polynomial model **BBL: of what order?** provided the best fit. FACF generally increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which they started to decrease (Figs. 4, S4-S5). The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, $R^2$ = 0.33. Data above this point were not available, but the model trend suggested a saturation point around 5000mm MAP.

There was a significant additive effect of MAT and MAP on $GPP$, $ANPP$ and $R_{auto}$ (Fig. 3, Table S3). Accounting for MAT, MAP had a substantial positive effect on $GPP$ and $R_{auto}$ and a small negative effect on $ANPP$. There was a significant interactive effect between MAT and MAP for $NPP$ and $ANPP_{stem}$ (Fig. 3, Table S3). The interaction was negative for $NPP$ and positive for $ANPP_{stem}$. For $ANPP_{foliage}$, $BNPP$, $BNPP_{fine.root}$, and $R_{root}$, MAP did not have a signficant effect when accounting for MAT (Fig. 3, Table S3). For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model. {**need to confirm this given changes in MAT MAP results (or you could just drop the sentence.)**}
**BBL: this paragraph is pretty dense. Perhaps distill to a summary sentence and move rest to SI?**

@@

We found a significant relationship between C flux and potential evapotranspiration for all FACF. $ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (Fig. 4c, S4-5; Table S2). We found strong evidence for a saturation point or peak with PET: FACF tended to increase at values below 1000mm, before saturating between 1200 and 1700mm. There was also evidence that FACF begin to decrease at values above 1800mm PET.

@@

Temperature seasonality was a significant predictor of FACF. $GPP$, $NPP$, $ANPP$, and $R_{root}$ exhibited a polynomial relationship with seasonality (all p<0.05; Figs. 4e, S6-7; Table S2). $ANPP_{foliage}$, $ANPP_{stem}$ and $R_{auto}$ decreased linearly with temperature seasonality (all p<0.05; Figs. 4e, S6-S7; Table S2). Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of FACF (Table S2). FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower. **BBL: perhaps put this into an ecosystem context; what are these? Aseasonal subtropical places?**

In contrast, there was no significant effect of precipitation seasonality on FACF, and both maximum vapour pressure deficit, and water stress months were poor or non-significant explainers of variation in FACF (Figs. S6-S7; Table S2).

@@

When FACF were standardized by growing season length, correlations with growing season climate--including temperature, precipitation, solar radiation, and PET--were generally weak (Figs. S8-S9). Speficifally, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05). Similarly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05). Growing season solar radiation had a positive influence on $GPP$ ($R^2$ = 0.21, p<0.001), $NPP$ ($R^2$ = 0.21, p<0.001), $BNPP$ ($R^2$ = 0.16, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.12, p<0.01). Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.18, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.11, p<0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p<0.05). {**Becky, please verify/ edit the following:** There were no other signifcant correlations between growing season length-standardized FACF (9 variables in Table 2) and growing season climate (**which variables?**)}.

@@

These findings clarify how FACF vary with latitude and climate on a global scale. Past studies have differed in their conclusions regarding the relationship between FACF and latitude or its correlates (Table 1, *H1*)--quite possibly because of lack of standardization with respect to stand age and disturbance history. Our findings indicate that, among mature, undisturbed stands, FACF are unambiguously highest in the tropical regions, and the relationship is approximately linear (Fig. 2). This contrasts with the suggestion that productivity of temperate forests is similar to that of tropical forests [@huston_global_2009]. Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects. In addition, because carbon allocation varies with stand age [@delucia_forest_2007] (**Doughty et al. 2018: DOI: 10.1111/btp.12504; Anderson-Teixeira et al. 2013: doi: 10.1111/gcb.12194**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients. For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**Doughty et al. 2018: DOI: 10.1111/btp.12504**);[@delucia_forest_2007; @piao_forest_2010; @kunert_understanding_2019].

We show that FACF are broadly consistent in their responses to climate drivers on the global scale, with at most modest trends in C allocation among the variable pairs tested (Figs. 2, S3). 
This parallels the observation that C allocation across multiple FACF varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011], 
and is not surprising given that trees face universal constraints in carbon allocation (**REFS**). 
The one trend in allocation that emerged from our analysis was a tendency for greater C allocation aboveground--and to foliage in particular--in warm tropical climates (Fig. S3). This is consistent with observations of increasing allocation to roots with declining temperature across a tropical elevational gradient [@moser_elevation_201], 
and with observations and theory predicting relatively higher belowground carbon allocation at higher latitudes [@gill_belowground_2016].
It is also consistent with findings that as forest productivity increases, proportionaly more carbon is allocated to $ANPP_{foliage}$ relative to $BNPP_{fine.root}$ (*Chen et al. 2019; doi: 10.1111/ele.13193*) or $ANPP_{stem}$ [@hofhansl_new_2015]. **( check / comment on  @litton_carbon_2007 )**

One previously hypothesized trend that was not supported by our analysis was that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which is based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016]. Our analysis, limited to mature forests, shows no such trend (Fig. S3). $CUE$ is known to decline with forest age [@delucia_forest_2007; @piao_forest_2010; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007]. (*This finding may have some important implications for modeling and our broader understanding. I'm not sure offhand how much theory has been built on the idea of low CUE in tropical forests because of warm temperatures...*)

**drop this paragraph ([issue 78](https://github.com/forc-db/Global_Productivity/issues/78))?** 
One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$) than in subsidiary fluxes ($BNPP_{fine.root}$, $R_{root}$, $ANPP_{stem}$) (Fig. 2; Table S2). There are two, non-exclusive, potential explanations for this. First, it may be that methodological variation is larger relative to flux magnitude for some of the subsidiary fluxes. Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{fine.root}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{root}$), and sampling depth is variable and often insufficient to capture the full soil profile. $ANPP_{stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled (*Clark et al. 2001 Ecological Applications*). However, methodological variation and uncertainty affect all of fluxes considered here, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes. Second, differences among variables in the proportion of variation explained by climate may be attributable to more direct climatic control over $GPP$ than subsidiary fluxes. That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as many other local- and regional-scale factors (**REFS**).

The latitudinal gradient in FACF (Fig. 2) is driven primarily by temperature-related climate variables, the effects of which are moderated by moisture availability (Table 1, *H2-H3*; Figs. 3-4). Indeed, MAT and MAP have long been identified as primary global-scale drivers of FACF [*Lieth 1973: doi 10.1007/BF01536729*; **REFS**; @taylor_temperature_2017]. It is not appropriate to attempt to identify individual mean annual climate variables as mechanistic drivers of FACF because many climate variables co-vary across the latitudinal gradient (Fig. S2), because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means. However, it remains informative to consider these relationships. Among the temperature-related climate variables, $MAT$ is generally the most strongly correlated with FACF (Table S2)--perhaps in part because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables were extracted from global gridded data products, introducing spatial scaling errors.

@@

FACF are negatively correlated with temperature seasonality (Table 1, *H4*; Fig. 4e), and are minimal during cold- or dry- dormant seasons. To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *H5*; @kerkhoff_plant_2005; *Anderson et al. 2006: doi:10.1111/j.1461-0248.2006.00914.x *; *Enquist et al. 2007: doi: 10.1111/j.1365-2486.2006.01222.x*;  @michaletz_convergence_2014]. We found that the sort of simple metric needed to define growing season at a global scale [@kerkhoff_plant_2005] was rough **BBL: ?** with respect to temperature and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow. A temperature-defined growing season length had stong positive correlation with FACF (Fig. 4f), but explained less variation than $MAT$. Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosyntheiss or respiration by high-latitude forests (*Enquist et al. 2007: doi: 10.1111/j.1365-2486.2006.01222.x*). However, there remained a number of significant correlations with growing season climatic conditions, suggesting that climatic conditions remain influential within the growing season. We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains crude relative to the timescales on which climate affects plant metabolism, and does not advance statistical predictive power. Mechanistic accounting for climatic effects on global FACF patterns instead requires models representing physiologically meaningful timescales (*e.g.*, **refs**). **BBL: maybe cite e.g. ED2? http://dx.doi.org/10.5194/gmd-12-4309-2019 **

Our analysis clarifies how FACF vary with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming. Our findings show that higher temperatures with similar moisture availability result in a generalized acceleration of FACF (Figs. 2-3). This is consistent with observations of continental- to global-scale increases in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018]. However, increasing C flux rates are by no means universal [@rutishauser_testing_2020] (**MORE REFS**), likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances. Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapated and acclimatized will not necessarily parallel responses across across geographic gradients in climate. Nevertheless, as we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.

>>>>

All fluxes increased with MAT (all p<0.05; Figs. 3-4, S4-S5, Table S2). 
For eight of the nine fluxes, this relationship was linear.
For only one variable, $BNPP$, did a lognormal fit provide signficant improvement over a first-order linear relatinship. 
*As with latitude, MAT tended to explain more variation in the larger flux ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).* **update this**

MAP was a significant (p<0.05) predictor of all fluxes (Figs. 4a, S4-S5; Table S2). 
However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 25% of variation in FACF. 
For five of the nine fluxes, a second-order polynomial model provided the best fit, while first-order linear and lognormal relationships provided each provided the best fit for two flxues. 
All fluxes increased with MAP up to at least 2000 mm, above which responses were variable (Figs. 4, S4-S5).

There was a significant additive effect of MAT and MAP on $GPP$, $ANPP$ and $R_{auto}$ (Fig. 3, Table S3). Accounting for MAT, MAP had a substantial positive effect on $GPP$ and $R_{auto}$ and a small negative effect on $ANPP$. There was a significant interaction between MAT and MAP for $NPP$ and $ANPP_{stem}$ (Fig. 3, Table S3). The interaction was negative for $NPP$ and positive for $ANPP_{stem}$. For $ANPP_{foliage}$, $BNPP$, $BNPP_{fine.root}$, and $R_{root}$, MAP did not have a signficant effect when accounting for MAT (Fig. 3, Table S3). For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model. {**need to confirm this given changes in MAT MAP results (or you could just drop the sentence.)**}
**BBL: this paragraph is pretty dense. Perhaps distill to a summary sentence and move rest to SI?**

@@

We found a significant relationship between C flux and potential evapotranspiration for all FACF. $ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (Fig. 4c, S4-5; Table S2). We found strong evidence for a saturation point or peak with PET: FACF tended to increase at values below 1000mm, before saturating between 1200 and 1700mm. There was also evidence that some FACF begin to decrease at values above 1800mm PET.

@@

Temperature seasonality was, overall, the best predictor of FACF. **(update this paragraph based on new results)**
$GPP$, $NPP$, $ANPP$, and $R_{root}$ exhibited a polynomial relationship with seasonality (all p<0.05; Figs. 4e, S6-7; Table S2). $ANPP_{foliage}$, $ANPP_{stem}$ and $R_{auto}$ decreased linearly with temperature seasonality (all p<0.05; Figs. 4e, S6-S7; Table S2). Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of FACF (Table S2). 
FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower. **BBL: perhaps put this into an ecosystem context; what are these? Aseasonal subtropical places?**

In contrast, there was no significant effect of precipitation seasonality on FACF, and both maximum vapour pressure deficit, and water stress months were poor or non-significant predictors of variation in FACF (Figs. S6-S7; Table S2).

@@

When FACF were standardized by growing season length (in monthly increments), correlations with growing season climate--including temperature, precipitation, solar radiation, and PET--were generally weak (Figs. S8-S9). Speficifally, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05). Similarly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05). Growing season solar radiation had a positive influence on $GPP$ ($R^2$ = 0.21, p<0.001), $NPP$ ($R^2$ = 0.21, p<0.001), $BNPP$ ($R^2$ = 0.16, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.12, p<0.01). Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.18, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.11, p<0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p<0.05). {**Becky, please verify/ edit the following:** There were no other signifcant correlations between growing season length-standardized FACF (9 variables in Table 2) and growing season climate (**which variables?**)}.

@@

These findings clarify how FACF vary with latitude and climate on a global scale. Past studies have differed in their conclusions regarding the relationship between FACF and latitude or its correlates (Table 1, *H1*)--quite possibly because of lack of standardization with respect to stand age and disturbance history. Our findings indicate that, among mature, undisturbed stands, FACF are unambiguously highest in the tropical regions, and the relationship with latitude is approximately linear (Fig. 2). This contrasts with the suggestion that productivity of temperate forests is similar to that of tropical forests [@huston_global_2009]. Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects. In addition, because carbon allocation varies with stand age [@delucia_forest_2007] (**Doughty et al. 2018: DOI: 10.1111/btp.12504; Anderson-Teixeira et al. 2013: doi: 10.1111/gcb.12194**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients. For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**Doughty et al. 2018: DOI: 10.1111/btp.12504**);[@delucia_forest_2007; @piao_forest_2010; @kunert_understanding_2019].

We show that FACF are broadly consistent in their responses to climate drivers on the global scale, with at most modest trends in C allocation among the variable pairs tested (Figs. 2, S3). 
This parallels the observation that C allocation across multiple FACF varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011], 
and is not surprising given that trees face universal constraints in carbon allocation (**REFS**). 
The one trend in allocation that emerged from our analysis was a tendency for greater C allocation aboveground--and to foliage in particular--in warm tropical climates (Fig. S3). This is consistent with observations of increasing allocation to roots with declining temperature across a tropical elevational gradient [@moser_elevation_2011], 
and with observations and theory predicting relatively higher belowground carbon allocation at higher latitudes [@gill_belowground_2016].
It is also consistent with findings that as forest productivity increases, proportionaly more carbon is allocated to $ANPP_{foliage}$ relative to $BNPP_{fine.root}$ (*Chen et al. 2019; doi: 10.1111/ele.13193*) or $ANPP_{stem}$ [@hofhansl_new_2015]. **( check / comment on  @litton_carbon_2007 )**

One previously hypothesized trend that was not supported by our analysis was that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which is based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016]. Our analysis, limited to mature forests, shows no such trend (Fig. S3). $CUE$ is known to decline with forest age [@delucia_forest_2007; @piao_forest_2010; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007]. (*This finding may have some important implications for modeling and our broader understanding. I'm not sure offhand how much theory has been built on the idea of low CUE in tropical forests because of warm temperatures...Helene is not aware of any.*)

**update this paragraph ([issue 78](https://github.com/forc-db/Global_Productivity/issues/78))** 
One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$) than in subsidiary fluxes ($BNPP_{fine.root}$, $R_{root}$, $ANPP_{stem}$) as quantified by R^2^ (Fig. 2; Table S2). There are two, non-exclusive, potential explanations for this. First, it may be that methodological variation is larger relative to flux magnitude for some of the subsidiary fluxes. Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may use fundamentally different approaches in different sites (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{fine.root}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{root}$), and sampling depth is variable and often insufficient to capture the full soil profile. $ANPP_{stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled (*Clark et al. 2001 Ecological Applications*). However, methodological variation and uncertainty affect all of fluxes considered here, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes. Second, differences among variables in the proportion of variation explained by climate may be attributable to more direct climatic control over $GPP$ than subsidiary fluxes. That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as many other local- and regional-scale factors (**REFS**).

The latitudinal gradient in FACF (Fig. 2) is driven primarily by temperature-related climate variables, the effects of which are moderated by moisture availability (Table 1, *H2-H3*; Figs. 3-4).  MAT and MAP have long been identified as primary global-scale drivers of FACF [*Lieth 1973: doi 10.1007/BF01536729*; **REFS**; @taylor_temperature_2017]. It is not appropriate to attempt to identify individual mean annual climate variables as mechanistic drivers of FACF because many climate variables co-vary across the latitudinal gradient (Fig. S2), because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means. However, it remains informative to consider these relationships. Among the temperature-related climate variables, $MAT$ is generally the most strongly correlated with FACF (Table S2)--perhaps in part because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables were extracted from global gridded data products, introducing spatial scaling errors.

@@

FACF are negatively correlated with temperature seasonality (Table 1, *H4*; Fig. 4e), and are minimal during cold- or dry- dormant seasons. To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *H5*; @kerkhoff_plant_2005; *Anderson et al. 2006: doi:10.1111/j.1461-0248.2006.00914.x *; *Enquist et al. 2007: doi: 10.1111/j.1365-2486.2006.01222.x*;  @michaletz_convergence_2014]. We found that the sort of simple metric needed to define growing season at a global scale [@kerkhoff_plant_2005] was rough **BBL: ?** with respect to temperature and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow. A temperature-defined growing season length had stong positive correlation with FACF (Fig. 4f), but explained less variation than $MAT$. Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests (*Enquist et al. 2007: doi: 10.1111/j.1365-2486.2006.01222.x*). However, there remained a number of significant correlations with growing season climatic conditions, suggesting that climatic conditions remain influential within the growing season. We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains crude relative to the timescales on which climate affects plant metabolism, and does not advance statistical predictive power. Mechanistic accounting for climatic effects on global FACF patterns instead requires models representing physiologically meaningful timescales (*e.g.*, **refs**). **BBL: maybe cite e.g. ED2? http://dx.doi.org/10.5194/gmd-12-4309-2019 **

Our analysis clarifies how FACF vary with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming. Our findings show that higher temperatures with similar moisture availability result in a generalized acceleration of FACF (Figs. 2-3). This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018]. However, increasing C flux rates are by no means universal [@rutishauser_testing_2020] (**MORE REFS**), likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances. **(discuss new Sullivan paper that finds higher max temperatures associated to lower ANPP_stem)**
Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapated and acclimatized will not necessarily parallel responses across geographic gradients in climate. Nevertheless, as we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.
